---
title: "Interactive Charts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive Charts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
library(ggplot2)
library(plotly)

# Load the small pre-aggregated data shipped with the package. Try the installed
# package first; fall back to the source tree (used when the site is built).
.path <- system.file("extdata", "explore_data.rds", package = "VietnamWarData")
if (!nzchar(.path) || !file.exists(.path)) .path <- "../inst/extdata/explore_data.rds"
ex <- readRDS(.path)
```

These charts are interactive and run entirely in your browser — no R session
required. Drag the year slider (or press play) to watch patterns change over
time; hover for exact values, and click-drag to zoom. The underlying figures
come from the combined incident database (`get_comb_inc_dta()`), pre-aggregated
for use here.

## Ground combat charts {.tabset .tabset-pills}

### Incident density over time

A 2-D density heat map of ground combat incidents across South Vietnam,
1963–1973, split by initiating side. Use the slider to step through the years.

```{r density, out.width = "100%", fig.height = 6}
g <- ggplot(ex$density_grid, aes(lon_bin, lat_bin, fill = n, frame = year)) +
  geom_tile() +
  facet_wrap(~ aggressor_side) +
  scale_fill_viridis_c(option = "inferno", trans = "log10", name = "Incidents") +
  coord_fixed(expand = FALSE) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 12) +
  theme(axis.text = element_blank(), panel.grid = element_blank(),
        strip.text = element_text(face = "bold"))

ggplotly(g, height = 560) |>
  animation_opts(frame = 650, transition = 0, redraw = TRUE) |>
  animation_slider(currentvalue = list(prefix = "Year "))
```

### Event counts

Enemy incidents by action category — press play or drag the slider to see how
the mix of harassment, terrorism, mining, and attacks shifted year to year.

```{r counts-cat, out.width = "100%", fig.height = 5}
plot_ly(
  ex$enemy_cat_year,
  x = ~n, y = ~enemy_action_category, frame = ~year,
  type = "bar", orientation = "h",
  marker = list(color = "#7c1f12"),
  hovertemplate = "%{y}: %{x:,} incidents<extra></extra>"
) |>
  layout(
    xaxis = list(title = "Number of incidents"),
    yaxis = list(title = "", categoryorder = "total ascending"),
    margin = list(l = 130)
  ) |>
  animation_opts(frame = 700) |>
  animation_slider(currentvalue = list(prefix = "Year "))
```

Incidents per month — hover for monthly totals; click-drag to zoom into a
period (e.g. the 1968 Tet Offensive).

```{r counts-ts, out.width = "100%", fig.height = 4.5}
g_ts <- ggplot(ex$by_month, aes(ym, n, color = aggressor_side)) +
  geom_line(linewidth = 0.6) +
  scale_color_brewer(palette = "Set1", name = NULL) +
  labs(x = NULL, y = "Incidents / month") +
  theme_minimal(base_size = 12)

ggplotly(g_ts, height = 380)
```

## {-}

Want to build these yourself? See the
[satellite](satellite-maps.html) and [counts](incident-counts.html) vignettes
for the underlying (static) recipes, or grab the data with
`get_comb_inc_dta()`.
